---
title: "STA-230 Midterm 3 REDO"
author: "Finn Rowles, Jafar Jarrar, Delaney Nguyen"
date: "2025-04-26"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
library(odbc)
library(RMySQL)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dbplyr)
library(tidyr)
```

```{r}
# Import the data from files.
data <- read.csv("/Users/finnrowles/Documents/2025SpringSTA230/train.csv")
data <- data %>% na.omit # This omits any N/A observations. This only omits ~11
# orders from burlington vermont due to an N/A postal code. Won't effect the
# general integrity of our data.
```

```{r}
colnames(data)[colnames(data) == "Row.ID"] = "RowID"
colnames(data)[colnames(data) == "Order.ID"] = "OrderID"
colnames(data)[colnames(data) == "Customer.ID"] = "CustomerID"
colnames(data)[colnames(data) == "Order.Date"] = "OrderDate"
colnames(data)[colnames(data) == "Ship.Date"] = "ShipDate"
colnames(data)[colnames(data) == "Customer.Name"] = "CustomerName"
colnames(data)[colnames(data) == "Sub.Category"] = "SubCategory"
colnames(data)[colnames(data) == "Product.Name"] = "ProductName"
colnames(data)[colnames(data) == "Ship.Mode"] = "ShipMode"
colnames(data)[colnames(data) == "Postal.Code"] = "PostalCode"
colnames(data)[colnames(data) == "Product.ID"] = "ProductID"

colnames(data)

# Motivation for our primary/foreign keys : identify distinct
# observations in rows - what describes other variables?
n_distinct(data$Order.ID)

n_distinct(data$Customer.ID)

n_distinct(data$Postal.Code)

n_distinct(data$City)
```
# Table 1:
Primary key: row id
Variables of interest: "Order.ID" , "Order.Date" , "Ship.Date" , "Ship.Mode", Sales 
Foreign keys: "Customer.ID", "Product.ID"

# Table 2:
Primary key: product_id
Variables of interest: "product_name", "product_category", "Category", "Sub.Category"

# Table 3:
Primary key: customer_id
Variables of interest: "customer_name", customer_age", "gender"

# Sales : quantity * price for a transaction?

```{r}
midterm3 = dbConnect(RSQLite::SQLite(), ":memory:")

dbWriteTable(midterm3, "data", data)

dbExecute(midterm3, "CREATE TABLE locationInfo (
          PostalCode INTEGER PRIMARY KEY NOT NULL,
          City varchar(64) NOT NULL, 
          State varchar(20) NOT NULL,
          Region varchar(20) NOT NULL)")

dbExecute(midterm3, "CREATE TABLE productInfo (
          ProductID varchar(26) PRIMARY KEY NOT NULL,
          Category varchar(20) NOT NULL, 
          SubCategory varchar(20) NOT NULL,
          ProductName varchar(128) NOT NULL)")

dbExecute(midterm3, "INSERT INTO locationInfo (PostalCode, City, State, Region)
  SELECT PostalCode, City, State, Region
  FROM data
  GROUP BY PostalCode")

dbExecute(midterm3, "INSERT INTO productInfo (ProductID, Category, SubCategory, ProductName)
  SELECT ProductID, Category, SubCategory, ProductName
  FROM data
  GROUP BY ProductID")

dbGetQuery(midterm3, "SELECT * from productInfo")

# Credit to W3schools SQL documentation for understanding declaring foreign
# keys.
dbExecute(midterm3, "CREATE TABLE primaryTable (
          RowID INTEGER PRIMARY KEY NOT NULL,
          Sales REAL NOT NULL,
          FOREIGN KEY (CustomerID) REFERENCES customerInfo(CustomerID),
          FOREIGN KEY (OrderID) REFERENCES shippingInfo(OrderID)
          FOREIGN KEY (PostalCode) REFERENCES locationInfo(PostalCode),
          FOREIGN KEY (ProductID) REFERENCES productInfo(ProductID)")
```
